Replication Script to JSS Submission

### 2016-05-20 JSS/jss-style/article.tex
replication script

\section{Available Software}

@@

The \proglang{R} package \proglang{pedigree} is named after the standardized chart used to study human family lines, and sometimes used to select breeding of animals, such as show dogs (\citealt{ped}). This package does provide tools that perform methods on parent-child datasets, such as rapidly determining the generation count for each member in the pedigree. However, it does not provide any visualization tools.

Another \proglang{R} package called \proglang{kinship2} does produce basic pedigree charts (\citealt{kin}). In Figure \ref{fig:kinshipFig}, we provide an example pedigree chart from the \proglang{kinship2} package vignette. This pedigree chart adheres to the standard set of symbols used for visualizing genealogical structures: Males are represented with squares and females with circles. Parents are connected to each other by horizontal lines, and to their children by vertical lines. Siblings are connected by horizontal sibship lines. Even though this standard pedigree chart creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational breeding occurs, as is often the case in agronomic genealogical lineages.

We demonstrate how the standardized pedigree charts in the \proglang{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational breeding case in Figure \ref{fig:kinshipFig}. In that figure, each generation is defined by its position on the vertical axis, with the first generation containing individuals 201 and 202. We superimposed green-highlighted individual 215 onto the pedigree chart for explanatory purposes. Its parents are individuals 201 and 206, which are from generations one and two, respectively, and have a parent-child relationship between themselves. As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual. Hence, individual 215 should be displayed in both second and third generational positions on the vertical axis. However, standard pedigree tools only allow for an individual to be displayed once. As a result, in special cases where inter-generational breading occurs, such as in agronomic applications, standardized tools for visualizing genealogical information ambiguously portray the genealogical dataset.

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.65\textwidth]{kinshipFig}}
    \caption{Example pedigree chart from the \proglang{kinship2} package, where the vertical axis denotes generation count. We superimposed green-highlighted individual 215 for explanatory purposes. As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual. Hence, it should be displayed twice on the vertical axis, once for each of its generation counts. However, standard pedigree tools only allow for an individual to be displayed once, leading to ambiguous portrayal of the genealogical dataset.}
    \label{fig:kinshipFig}
\end{figure}

In addition, popular graph drawing software such as \proglang{GraphViz} and \proglang{Cytoscape} can be used to visualize genealogical structures (\citealt{graphvizCit}, \citealt{cytoscapeCit}). Graphs are defined as objects with sets of nodes and edges, where sets indicate that their comprised elements cannot be repeated. In other words, graphical structures do not allow for repeated nodes, and hence, as is the case with the aforementioned \proglang{R} packages, these popular graph plotting software cannot precisely portray the genealogical dataset in cases of inter-generational breeding.

We again illustrate this problem in Figure \ref{fig:Graph} with an example genealogy using popular graph drawing software like \proglang{GraphViz} and \proglang{Cytoscape}. Here, generation count is denoted by the vertical axis. As was shown in Figure \ref{fig:kinshipFig}, here too we superimpose a green node that has parents from two different generations. This green node is both a second and third generation individual, and should be displayed in both corresponding generation positions on the vertical axis. However, standard graph visualization tools only allow for a given node to be displayed once. As a result, this green node must be ambiguously positioned in either the second or third generation position; in the figure, it is denoted as a third generation individual. In Section \ref{remedy}, we will demonstrate \pkg{ggenealogy} plots that can remedy these problems.

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.6\textwidth]{Graph}}
    \caption{Example genealogical display using popular graph software like \proglang{GraphViz} and \proglang{Cytoscape}, with generation count denoted by the vertical axis. As was shown in Figure \ref{fig:kinshipFig}, the green node has parents from two different generations, and hence must be ambiguously positioned as one of two generation counts.}
    \label{fig:Graph}
\end{figure}

\section{Example Datasets}
\label{exData}

@@

\subsection{Soybean Genealogy}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("ggenealogy")
R> library("ggenealogy")
R> data(sbGeneal)
R> str(sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
'data.frame':	390 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ year        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{CodeOutput}
\end{CodeChunk}

\subsection{Academic Genealogy of Statisticians}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data(statGeneal)
R> dim(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 8165    6
\end{CodeOutput}
\begin{CodeInput}
R> colnames(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] "child"   "parent"  "year"    "country" "school"  "thesis"
\end{CodeOutput}
\end{CodeChunk}

\section{Genealogical Input Format}

@@

\section{Generating a Graphical Object}

@@

\section{Plotting a Shortest Path}

@@

\section{Superimposing Shortest Path on Tree}

@@

\section{Plotting Ancestors and Descendants by Generation}
\label{remedy}

@@

\section{Plotting Distance Matrix}

@@

\section{Academic Genealogy of Statisticians}

@@

\begin{CodeChunk}
\begin{CodeInput}
>R library(dplyr)
>R indVec <- getNodes(statGeneal)
>R indVec <- indVec[which(indVec != "", )]
>R dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
>R numDesc <- sapply(indVec, dFunc)
>R table(numDesc)
\end{CodeInput}
\begin{CodeOutput}
numDesc
   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14 
6252  322  145   88   58   36   31   22   23   14   17   13   14    9    9 
  15   16   17   18   19   20   21   22   23   24   25   26   27   29   30 
   6    4    3    2    5    7    5    3    3    2    2    6    1    1    3 
  34   37   38   40   41   44   45   48   49   61   62   75   77   84  159 
   2    1    1    1    1    1    1    1    2    1    1    1    1    1    1
\end{CodeOutput}
\begin{CodeInput}
R> which(numDesc == 159)
\end{CodeInput}
\begin{CodeOutput}
David Cox 
     1980
\end{CodeOutput}
\end{CodeChunk}

@@

We could still improve upon Figure \ref{fig:plotCBNoText}. Even though we may be primarily interested in understanding how the path of interest is overlaid across the entire genealogical structure, we could, upon viewing the entire structure, also develop an interest in nodes that are not on the path of interest but are revealed to stand out among the rest of the genealogical structure. For instance, in Figure \ref{fig:plotCBNoText}, it may be of interest for us to determine the names of the few individuals who obtained their statistics degrees before 1900. Fortunately, within the \code{plotPathOnAll()} function, there is a variable \code{animate} that we can set to a value of TRUE to create an interactive version of the figure that allows us to hover over individual illegible labels and immediately receive their labels in a readable format. A short video demonstration of these interactive features can be viewed below.

@@

\section{Future Avenues}

>>>>

\section{Available software}

@@

The \proglang{R} package \pkg{pedigree} is named after the standardized chart used to study human family lines, and sometimes used to select breeding of animals, such as show dogs (\citealt{ped}). This package does provide tools that perform methods on parent-child datasets, such as rapidly determining the generation count for each member in the pedigree. However, it does not provide any visualization tools.

Another \proglang{R} package called \pkg{kinship2} does produce basic pedigree charts (\citealt{kin}). In Figure \ref{fig:kinshipFig}, we provide an example pedigree chart from the \pkg{kinship2} package vignette. This pedigree chart adheres to the standard set of symbols used for visualizing genealogical structures: Males are represented with squares and females with circles. Parents are connected to each other by horizontal lines, and to their children by vertical lines. Siblings are connected by horizontal sibship lines. Even though this standard pedigree chart creates powerful charts that can be applied across many applications, it cannot provide unequivocal information in situations where inter-generational breeding occurs, as is often the case in agronomic genealogical lineages.

We demonstrate how the standardized pedigree charts in the \pkg{kinship2} package generate ambiguous results in such scenarios by superimposing a hypothetical inter-generational breeding case in Figure \ref{fig:kinshipFig}. In that figure, each generation is defined by its position on the vertical axis, with the first generation containing individuals 201 and 202. We superimposed green-highlighted individual 215 onto the pedigree chart for explanatory purposes. Its parents are individuals 201 and 206, which are from generations one and two, respectively, and have a parent-child relationship between themselves. As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual. Hence, individual 215 should be displayed in both second and third generational positions on the vertical axis. However, standard pedigree tools only allow for an individual to be displayed once. As a result, in special cases where inter-generational breading occurs, such as in agronomic applications, standardized tools for visualizing genealogical information ambiguously portray the genealogical dataset.

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.65\textwidth]{kinshipFig}}
    \caption{Example pedigree chart from the kinship2 package, where the vertical axis denotes generation count. We superimposed green-highlighted individual 215 for explanatory purposes. As an offspring of a parent-child relationship, individual 215 is both a second and third generation individual. Hence, it should be displayed twice on the vertical axis, once for each of its generation counts. However, standard pedigree tools only allow for an individual to be displayed once, leading to ambiguous portrayal of the genealogical dataset.}
    \label{fig:kinshipFig}
\end{figure}

In addition, popular graph drawing software such as \pkg{GraphViz} and \pkg{Cytoscape} can be used to visualize genealogical structures (\citealt{graphvizCit}, \citealt{cytoscapeCit}). Graphs are defined as objects with sets of nodes and edges, where sets indicate that their comprised elements cannot be repeated. In other words, graphical structures do not allow for repeated nodes, and hence, as is the case with the aforementioned \proglang{R} packages, these popular graph plotting software cannot precisely portray the genealogical dataset in cases of inter-generational breeding.

We again illustrate this problem in Figure \ref{fig:Graph} with an example genealogy using popular graph drawing software like \pkg{GraphViz} and \pkg{Cytoscape}. Here, generation count is denoted by the vertical axis. As was shown in Figure \ref{fig:kinshipFig}, here too we superimpose a green node that has parents from two different generations. This green node is both a second and third generation individual, and should be displayed in both corresponding generation positions on the vertical axis. However, standard graph visualization tools only allow for a given node to be displayed once. As a result, this green node must be ambiguously positioned in either the second or third generation position; in the figure, it is denoted as a third generation individual. In Section \ref{remedy}, we will demonstrate \pkg{ggenealogy} plots that can remedy these problems.

\begin{figure}[H]
    \centering
    \fbox{\includegraphics[width=0.6\textwidth]{Graph}}
    \caption{Example genealogical display using popular graph software like \pkg{GraphViz} and \pkg{Cytoscape}, with generation count denoted by the vertical axis. As was shown in Figure \ref{fig:kinshipFig}, the green node has parents from two different generations, and hence must be ambiguously positioned as one of two generation counts.}
    \label{fig:Graph}
\end{figure}

\section{Example datasets}
\label{exData}

@@

\subsection{Soybean genealogy}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> install.packages("ggenealogy")
R> library("ggenealogy")
R> data("sbGeneal")
R> str(sbGeneal)
\end{CodeInput}
\begin{CodeOutput}
'data.frame':	390 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ year        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{CodeOutput}
\end{CodeChunk}

\subsection{Academic genealogy of statisticians}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data("statGeneal")
R> dim(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] 8165    6
\end{CodeOutput}
\begin{CodeInput}
R> colnames(statGeneal)
\end{CodeInput}
\begin{CodeOutput}
[1] "child"   "parent"  "year"    "country" "school"  "thesis"
\end{CodeOutput}
\end{CodeChunk}

\section{Genealogical input format}

@@

\section{Generating a graphical object}

@@

\section{Plotting a shortest path}

@@

\section{Superimposing shortest path on tree}

@@

\section{Plotting ancestors and descendants by generation}
\label{remedy}

@@

\section{Plotting distance matrix}

@@

\section{Academic genealogy of statisticians}

@@

\begin{CodeChunk}
\begin{CodeInput}
>R library("dplyr")
>R indVec <- getNodes(statGeneal)
>R indVec <- indVec[which(indVec != "", )]
>R dFunc <- function(var) nrow(getDescendants(var, statGeneal, gen = 100))
>R numDesc <- sapply(indVec, dFunc)
>R table(numDesc)
\end{CodeInput}
\begin{CodeOutput}
numDesc
   0    1    2    3    4    5    6    7    8    9   10   11   12   13   14 
6252  322  145   88   58   36   31   22   23   14   17   13   14    9    9 
  15   16   17   18   19   20   21   22   23   24   25   26   27   29   30 
   6    4    3    2    5    7    5    3    3    2    2    6    1    1    3 
  34   37   38   40   41   44   45   48   49   61   62   75   77   84  159 
   2    1    1    1    1    1    1    1    2    1    1    1    1    1    1
\end{CodeOutput}
\begin{CodeInput}
R> which(numDesc == 159)
\end{CodeInput}
\begin{CodeOutput}
David Cox 
     1980
\end{CodeOutput}
\end{CodeChunk}

@@

We could still improve upon Figure \ref{fig:plotCBNoText}. Even though we may be primarily interested in understanding how the path of interest is overlaid across the entire genealogical structure, we could, upon viewing the entire structure, also develop an interest in nodes that are not on the path of interest but are revealed to stand out among the rest of the genealogical structure. For instance, in Figure \ref{fig:plotCBNoText}, it may be of interest for us to determine the names of the few individuals who obtained their statistics degrees before 1900. Fortunately, within the \code{plotPathOnAll()} function, there is a variable \code{animate} that we can set to a value of TRUE to create an interactive version of the figure that allows us to hover over individual illegible labels and immediately receive their labels in a readable format. This interactive functionality comes from methods in the \pkg{plotly} package (\citealt{plotly}). A short video demonstration of these interactive features can be viewed below.

@@

\section{Future avenues}

### 2016-05-20 JSS/jss-style/replicationScript.Rnw
replication script

\documentclass{article}
\parindent 0pt
\title{Replication Script to JSS Submission:\\
\textsl{ggenealogy: An R Package for Visualizing Genealogical Data}
\author{Lindsay Rutter, Susan VanderPlas, Di Cook}
\date{\today}
\begin{document}
\maketitle
\SweaveOpts{concordance=TRUE}
\tableofcontents
<<>>=
install.packages("ggenealogy")
library("ggenealogy")
@
%%%%%%%%%%%%%%%
%% Section 1 %%
%%%%%%%%%%%%%%%
\section{Generating a Graphical Object}

Examine structure of sbGeneal example dataset:

<<>>=
data(sbGeneal)
str(sbGeneal)
@

@@

<<>>=
library("mstate")
tra2 <- !is.na(transMat(list(c(2), c(1))))
tra2
@
Use \texttt{mplskeleton} to construct a skeleton of the model parameter list \texttt{mpl}:
<<>>=
mpl <- mplskeleton(tmat = tra2)
str(mpl)
@
Parametrize the model and complete \texttt{mpl}:
<<>>=
mpl[[1]]$bhr[[2]] <- mpl[[2]]$bhr[[1]] <- function(t){return(0.5)}
## time-varying x2 and time-varying f(x2):
mpl[[1]]$eta[[2]] <- function(x.i, t){
  ifelse(t < 5, 
         return(1.0 * x.i[1] + 0.5 * x.i[2]), 
         return(1.0 * x.i[1] + 1.0 * x.i[3]))}
## time-varying x2 and time-varying f(x1):
mpl[[2]]$eta[[1]] <- function(x.i, t){
  ifelse(t < 5, 
         return(-0.5 * x.i[1] + 0.5 * x.i[2]), 
         return( 1.0 * x.i[1] + 0.5 * x.i[3]))}
@
Simulate event histories:
<<>>=
set.seed(123)
N <- 500
X <- matrix(nrow = N, ncol = 2, rnorm(2 * N))
X <- cbind(X, X[, 2] + runif(N)/10)
colnames(X) <- c("x1", "x2.t1", "x2.t2")
head(X)
Xstruc <- matrix(ncol = 2, nrow = 2, data = 0)
rownames(Xstruc) <- c("t1", "t2")
colnames(Xstruc) <- c("x1", "x2")
Xstruc[, 1] <- 1
Xstruc[, 2] <- c(2, 3)
Xstruc
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = c(5), Xstruc = Xstruc)
head(d$msm.basics)
head(d$ttsce[, 1:7])
head(d$tt.indicators)
@
Nelson-Aalen estimator of the cumulative hazard rate functions:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d = d$msm.basics, mpl = mpl, xlim = c(0, 15))
abline(a = 0, b = 0.5, col = "darkgrey")
@

Stratified partial likelihood modelling:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
m <- coxph(Surv(entry, exit, delta) ~ x1.12.t1 + x2.12.t1 + x1.12.t2 + 
             x2.12.t2 + x1.21.t1 + x2.21.t1 + x1.21.t2 + x2.21.t2 + 
             strata(trans), data = data.frame(d$msm.basics, d$ttsce))
round(summary(m)$coef[, c(1, 5)], 5)
plotcph(m)
points(c(c(1, 0.5, 1 - 1, 1 - 0.5, -0.5, 0.5, 1 - (-0.5), 0.5 - 0.5)), 
       pch = 16, col = "darkgrey")
@

Breslow estimate for cumulative hazard:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=be>>=
plotbe(m, mpl, xlim = c(0, 15), ylim = c(0, 10))
abline(a = 0, b = 0.5, col = "darkgrey")
@
%%%%%%%%%%%%%%%
%% Section 2 %%
%%%%%%%%%%%%%%%
\section{Illness-death Model with Recovery, a Non-linear Baseline Hazarad Rate Function and Non-linear Covariate Effects}
Prepare the simulation of event histories:
<<>>=
library("simMSM")
traIDM <- matrix(nrow = 3, ncol = 3, FALSE)
traIDM[1, 2] <- traIDM[1, 3] <- traIDM[2, 1] <- traIDM[2, 3] <- TRUE
## Or use mstate:
library("mstate")
traIDM <- !is.na(transMat(list(c(2, 3), c(1, 3), c())))
traIDM
## Build up mpl:
mpl <- mplskeleton(tmat = traIDM)
mpl[[1]]$bhr[[2]] <- function(t){0.5 + 0.4 * sin(t)}
mpl[[1]]$bhr[[3]] <- mpl[[2]]$bhr[[1]] <- 
  mpl[[2]]$bhr[[3]] <- function(t){0.5}
mpl[[1]]$eta[[2]] <- function(x.i, t){ 3 * (x.i^2)}
mpl[[1]]$eta[[3]] <- function(x.i, t){-3 * (x.i^2)}
mpl[[2]]$eta[[1]] <- mpl[[2]]$eta[[3]] <- 
  function(x.i, t){0.5 * x.i}
## simulate design matrix and event history data:
set.seed(123)
N <- 500
X <- matrix(ncol = 1, nrow = N, data = runif(N) - 0.5)
colnames(X) <- c("x")
Xstruc <- matrix(ncol = 1, nrow = 1, data = 1)
rownames(Xstruc) <- c("t1")
colnames(Xstruc) <- c("x")
Xstruc
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = NULL, Xstruc = Xstruc)
head(d$msm.basics)
@

Nelson-Aalen estimators:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d = d$msm.basics, mpl)
@

Convert to long data format and estimate a stratified Cox proportional hazards (partial likelihood) model:
<<>>=
ld <- tolongformat(d, mpl = mpl)
head(ld$msm.basics)
head(ld$tt.indicators)
m <- coxph(Surv(entry, exit, delta) ~ I(x.12^2) + I(x.13^2) + 
             x.21 + x.23 + strata(trans), 
           data = data.frame(ld$msm.basics, ld$ttsce))
round(summary(m)$coef[, c(1, 5)], 5)
@

Cox model coefficient illustration:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
plotcph(m)
points(c(3, -3, 0.5, 0.5), col = "darkgrey", pch = 16)
@

Breslow estimator:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotbe(m, mpl)
@

BayesX:
<<>>=
d <- data.frame(d$msm.basics, d$tt.indicators)
install.packages("BayesXsrc")
library("BayesXsrc")
p <- "delimiter = ;
dataset dat;
dat.infile using d.raw;
remlreg simmsm;
simmsm.mregress 
trans12=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
trans13=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
trans21=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2):
trans23=exit(baseline,gridchoice=all,nrknots=30,lambdastart=5)+x(psplinerw2),
family=multistate lefttrunc=entry state=from maxit=1000 using dat;
delimiter = return;"
write.table(d, row.names = FALSE, quote = FALSE, file = "d.raw")
writeLines(p, con = "prg")
run.bayesx(prg = "prg", verbose = FALSE)
@

Prepare plotting the results:
<<>>=
setwd("./output/")
log.bhr <- beta.fixed <- beta.x <- vector("list", 4)
log.bhr[[1]] <- read.table("simmsm_f_exit_logbaseline.res", header=T)
log.bhr[[2]] <- read.table("simmsm_f_2_exit_logbaseline.res", header=T)
log.bhr[[3]] <- read.table("simmsm_f_3_exit_logbaseline.res", header=T)
log.bhr[[4]] <- read.table("simmsm_f_4_exit_logbaseline.res", header=T)
beta.fixed[[1]] <- read.table("simmsm_FixedEffects.res", header=T)
beta.fixed[[2]] <- read.table("simmsm_FixedEffects_2.res", header=T)
beta.fixed[[3]] <- read.table("simmsm_FixedEffects_3.res", header=T)
beta.fixed[[4]] <- read.table("simmsm_FixedEffects_4.res", header=T)
beta.x[[1]] <- read.table("simmsm_f_x_pspline.res", header=T)
beta.x[[2]] <- read.table("simmsm_f_2_x_pspline.res", header=T)
beta.x[[3]] <- read.table("simmsm_f_3_x_pspline.res", header=T)
beta.x[[4]] <- read.table("simmsm_f_4_x_pspline.res", header=T)
setwd("../")
plot.bhr <- function(log.bhr, beta.fixed, t.max){
  index.for.plot <- c("12,0", "13,0", "21,0", "23,0")
  par(mfrow = c(2, 2), mar = c(3.1, 3.1, 0.1, 0.1), tcl = -0.3)
  for(q in 1:4){
    plot(1, 1, ylim=c(0, 3), xlim=c(0, t.max), 
         type = "n", xaxt = "n", yaxt = "n", 
         xlab = "", ylab = "", bty = "n")
    axis(1, at = c(0, 4, 8, 12), label = FALSE)
    mtext(side = 1, text = c(0, 4, 8, 12), at = c(0, 4, 8, 12), 
          line = 0.5, cex = 0.5)
    axis(2, label = FALSE, at = 0:2)
    mtext(side = 2, text = 0:2, at = 0:2, line = 0.5, cex = 0.5)
    mtext("t", side = 1, line = 1.4, cex = 0.7)
    mtext(bquote(paste(lambda[.(index.for.plot[q])],(t))), side = 2, 
          line = 1.4, cex = 0.7)
    hilf <- beta.fixed[[q]]
    if(q == 1){
      ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <- hilf$pmode[1] - 0.255
    }
    if(q == 2){
      ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <- hilf$pmode[1] + 0.255
    }
    if(q %in% c(3, 4)){
      intercept <- hilf$pmode[1]
    }
    hilf <- log.bhr[[q]]
    polygon(c(hilf$exit, rev(hilf$exit)), 
            exp(intercept + c(hilf$ci95lower, rev(hilf$ci95upper))), 
            border = "NA", col = "lightgrey")
    lines(hilf$exit, exp(hilf$pmode + intercept), col = 1, lwd = 2)
    time <- seq(0, t.max, length = 100)
    if(q == 1){lines(time, (0.5 + 0.4 * sin(time)), lwd = 2, lty = 2)}
    if(q == 2){lines(time, (0.5 + 0*time), lwd = 2, lty = 2)}
    if(q == 3){lines(time, 0.5 + 0*time, lwd = 2, lty = 2)}
    if(q == 4){lines(time, 0.5 + 0*time, lwd = 2, lty = 2)}
  }
}
plot.spline <- function(beta.x){
  index.for.plot <- c("12,x", "13,x", "21,x", "23,x")
  par(mfrow = c(2, 2), mar = c(3.1, 3.1, 0.1, 0.1), tcl = -0.3)
  for(q in 1:4){
    plot(1, 1, ylim = c(-1.2, 1.2), xlim=c(-0.5, 0.5), type = "n", 
         xaxt = "n", yaxt = "n", xlab = "", ylab = "", bty = "n")
    axis(1, at = c(-0.5, 0, 0.5), label = FALSE)
    mtext(side = 1, text = c(-0.5, 0, 0.5), at = c(-0.5, 0, 0.5), 
          line = 0.5, cex = 0.5)
    axis(2, label = FALSE, at = -1:1)
    mtext(side = 2, text = -1:1, at = -1:1, line = 0.5, cex = 0.5)
    mtext("x", side = 1, line = 1.4, cex = 0.7)
    mtext(bquote(f[.(index.for.plot[q])](x)), side = 2, 
          line = 1.4, cex = 0.7)
    if(q == 1){
       ## 0.255 = mean( 3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <-  0.255
    }
    if(q == 2){
      ## -0.255 = mean(-3 * (seq(-0.5, 0.5, length = 100)^2))
      intercept <- -0.255
    }
    if(q %in% c(3, 4)){
      intercept <- 0
    }
    hilf <- beta.x[[q]]
    polygon(c(hilf$x, rev(hilf$x)), 
            intercept + c(hilf$ci95lower, rev(hilf$ci95upper)), 
            border = "NA", col = "lightgrey")
    lines(hilf$x, hilf$pmode + intercept, col = 1, lwd = 2)
    x <- seq(-0.5, 0.5, length = 100)
    if(q == 1){lines(x, 3*(x^2), lwd = 2, lty = 2)}
    if(q == 2){lines(x,-3*(x^2), lwd = 2, lty = 2)}
    if(q == 3){lines(x, 0.5 * x, lwd = 2, lty = 2)}
    if(q == 4){lines(x, 0.5 * x, lwd = 2, lty = 2)}
  }
}
@

BayesX baseline hazard rate function estimates:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=bayesX_bhr>>=
plot.bhr(log.bhr, beta.fixed, t.max = print(max(d$exit)))
@

BayesX covariate effect function estimates:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=bayesX_x>>=
plot.spline(beta.x)
@
%%%%%%%%%%%%%%%
%% Section 3 %%
%%%%%%%%%%%%%%%
\section{Illness-death Model with Partial Markov Influences}
Model build up in a general way:
<<>>=
traIDM <- matrix(nrow = 3, ncol = 3, FALSE)
traIDM[1, 2] <- traIDM[1, 3] <- traIDM[2, 1] <- traIDM[2, 3] <- TRUE
mpl <- mplskeleton(tmat = traIDM)
mpl[[1]]$bhr[[2]] <- mpl[[1]]$bhr[[3]] <- mpl[[2]]$bhr[[1]] <- 
                     mpl[[2]]$bhr[[3]] <- function(t){0.25}
mpl[[1]]$eta[[2]] <- mpl[[1]]$eta[[3]] <- mpl[[2]]$eta[[1]] <- 
  mpl[[2]]$eta[[3]] <- function(x.i, t){
    ifelse(t < 5, 
      return(0.5 * x.i[1]), 
      return(0.5 * x.i[2]))}
set.seed(123)
N <- 500
X <- matrix(nrow = N, ncol = 1, rnorm(N))
X <- cbind(X, X[, 1] + rnorm(N)/10)
colnames(X) <- c("x1.t1", "x1.t2")
Xstruc <- matrix(ncol = 1, nrow = 2, data = 0)
rownames(Xstruc) <- c("t1", "t2")
colnames(Xstruc) <- c("x1")
Xstruc[, 1] <- c(1, 2)
Xstruc
@
Now set-up the partial Markov influences. 
Function \texttt{partial.markov.x} has to take 5 input arguments representig vectors of past history information. 
They have to take names \texttt{entry}, \texttt{exit}, \texttt{from}, \texttt{to}, and \texttt{delta}:
<<>>=
partial.markov.x <- function(entry, exit, from, to, delta){
  count.12 <- sum(as.numeric((from == 1) & (to == 2) & (delta == 1)))
  count.21 <- sum(as.numeric((from == 2) & (to == 1) & (delta == 1)))
  return(c(count.12, count.21))}
@
The object \texttt{partial.markov.eta} is a list of lists in analogy to \texttt{mpl}:
<<>>=
partial.markov.eta <- pmeskeleton(traIDM)
str(partial.markov.eta)
partial.markov.eta[[1]][[2]] <- function(x){return( 0.25 * x[1])}
partial.markov.eta[[1]][[3]] <- function(x){return( 0.5 * x[1])}
partial.markov.eta[[2]][[1]] <- function(x){return(-0.5 * x[1] + 0.25 * x[2])}
partial.markov.eta[[2]][[3]] <- function(x){return(0)}
@
Event history simulation:
<<>>=
d <- simeventhistories(n = N, mpl = mpl, X = X, max.time = 10, 
                       change.times = c(5), Xstruc = Xstruc, 
                       partial.markov.x = partial.markov.x, 
                       partial.markov.eta = partial.markov.eta)
@
Nelson-Aalen estimator of the cumulative hazard rate functions:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=nae>>=
plotnae(d$msm.basics, mpl = mpl)
abline(a = 0, b = 0.25, col = "darkgrey")
@

Stratified partial likelihood modelling:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=cph>>=
ld <- tolongformat(d = d, mpl = mpl)
head(ld$msm.basics, n = 10)
head(ld$ttsce[, 1:6], n = 10)
m <- coxph(Surv(entry, exit, delta) ~ x1 + 
             pm.x1.12 + pm.x1.13 + pm.x1.21 + pm.x2.21 + 
             strata(trans), 
           data = data.frame(ld$msm.basics, ld$ttsce))
plotcph(m)
points(c(0.5, 0.25, 0.5, -0.5, 0.25), pch = 16, col = "darkgrey")
@

Breslow estimate for cumulative hazard:
<<fig=TRUE, pdf=TRUE, height=5, width=7, label=be>>=
plotbe(m, mpl = mpl)
abline(a = 0, b = 0.25, col = "darkgrey")
@
\end{document}

>>>>

\documentclass{article}
\parindent 0pt
\title{Replication Script to JSS Submission}
\begin{document}

\author{Lindsay Rutter, Susan VanderPlas, Dianne Cook, Michelle A. Graham}
\date{\today}
\maketitle
\SweaveOpts{concordance=TRUE}
\tableofcontents

%%%%%%%%%%%%%%%
%% Section 3 %%
%%%%%%%%%%%%%%%
\section*{3. Example datasets}

First, we can load and examine the structure of the soybean genealogy example dataset (called sbGeneal):

<<>>=
rm(list=)
library("ggenealogy")
data("sbGeneal")
str(sbGeneal)
@

After that, we can load and examine the structure of the academic statistician genealogy example dataset (called statGeneal):

<<>>=
data("statGeneal")
dim(statGeneal)
colnames(statGeneal)
@

@@

\end{document}

### 2016-05-20 JSS/jss-style/replicationScript.tex
replication script

\documentclass{article}
\parindent 0pt
\title{Replication Script to JSS Submission:\\
\textsl{ggenealogy: An R Package for Visualizing Genealogical Data}
\author{Lindsay Rutter, Susan VanderPlas, Di Cook}
\date{\today}
\begin{document}
\maketitle
\SweaveOpts{concordance=TRUE}
\tableofcontents
<<>>=
install.packages("ggenealogy")
library("ggenealogy")
@
%%%%%%%%%%%%%%%
%% Section 1 %%
%%%%%%%%%%%%%%%
\section{Generating a Graphical Object}

Examine structure of sbGeneal example dataset:

<<>>=
data(sbGeneal)
str(sbGeneal)
@

>>>>

\documentclass{article}
\parindent 0pt
\title{Replication Script to JSS Submission}
\usepackage{Sweave}
\begin{document}

\author{Lindsay Rutter, Susan VanderPlas, Dianne Cook, Michelle A. Graham}
\date{\today}
\maketitle
\input{replicationScript-concordance}
\tableofcontents

%%%%%%%%%%%%%%%
%% Section 3 %%
%%%%%%%%%%%%%%%
\section*{3. Example datasets}

First, we can load and examine the structure of the soybean genealogy example dataset (called sbGeneal):

\begin{Schunk}
\begin{Sinput}
> rm(list=)
> library("ggenealogy")
> data("sbGeneal")
> str(sbGeneal)
\end{Sinput}
\begin{Soutput}
'data.frame':	390 obs. of  5 variables:
 $ child       : chr  "5601T" "Adams" "A.K." "A.K. (Harrow)" ...
 $ year        : num  1981 1948 1910 1912 1968 ...
 $ yield       : int  NA 2734 NA 2665 NA 2981 2887 2817 NA NA ...
 $ year.imputed: logi  TRUE FALSE TRUE FALSE FALSE FALSE ...
 $ parent      : chr  "Hutcheson" "Dunfield" NA "A.K." ...
\end{Soutput}
\end{Schunk}

After that, we can load and examine the structure of the academic statistician genealogy example dataset (called statGeneal):

\begin{Schunk}
\begin{Sinput}
> data("statGeneal")
> dim(statGeneal)
\end{Sinput}
\begin{Soutput}
[1] 8165    6
\end{Soutput}
\begin{Sinput}
> colnames(statGeneal)
\end{Sinput}
\begin{Soutput}
[1] "child"   "parent"  "year"    "country" "school"  "thesis" 
\end{Soutput}
\end{Schunk}
